---
title: "Credit Hours by Program — Streamlined"
author: "Tony Daza"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(DBI)
library(dplyr)
library(stringr)
library(odbc)
library(keyring)
library(tidyr)
library(ggplot2)
library(knitr)
library(kableExtra)
```

# Overview

This document calculates credit hours by program (subject) for tuition revenue
projections. Students are grouped into the following categories:

- **UGRD** — Traditional undergraduate, resident or non-resident
- **GRAD** — Graduate students, resident or non-resident (graduate-level courses only)
- **UFLEX** — University Flex students (identified by career code or term code prefix)
- **WI 60+ Audit** — Wisconsin residents aged 60+ who are auditing (no-load) — excluded from billed credit hours but tracked separately

Residency categories use `RES_TUITION_RES_LDESC`. Students previously coded
under the Midwest Student Exchange (which UW-Parkside no longer participates in)
are recoded to **Midwest Tuition Non-Resident**.

---

# Term Reference

```{r term-reference}
# ── Traditional terms (CLASS_TERM_CD begins with 1) ──────────────────────────
# 1257 = Summer 2024–2025
# 1261 = Fall   2025–2026
# 1263 = Winter 2025–2026
# 1265 = Spring 2025–2026
# 1267 = Summer 2025–2026

# ── UFlex terms (CLASS_TERM_CD begins with 5 or 6) ───────────────────────────
# 6506 = Jun–Aug 2025
# 6507 = Jul–Sep 2025
# 6508 = Aug–Oct 2025
# 6509 = Sep–Nov 2025
# 6510 = Oct–Dec 2025
# 6511 = Nov–Jan 2026
# 6512 = Dec–Feb 2026
# 6601 = Jan–Mar 2026
# 6602 = Feb–Apr 2026
# 6603 = Mar–May 2026

TRADITIONAL_TERMS <- c('1257','1261','1263','1265','1267')
UFLEX_TERMS       <- c('6506','6507','6508','6509','6510',
                        '6511','6512','6601','6602','6603')
ALL_TERMS         <- c(TRADITIONAL_TERMS, UFLEX_TERMS)
```


--- 

# Connect RDS 

```{r rds connect}

con_rds <- dbConnect(
  odbc::odbc(),
  dsn     = "Oracle RDS",
  uid     = "DAZA",
  pwd     = keyring::key_get("oracle_rds", "DAZA"),
  timeout = 10
)


```


# Connect CDR

```{r cdr connect}

# CDR
con_cdr <- dbConnect(
  odbc::odbc(),
  dsn = "Oracle CDR",          # DSN handles driver + TNS service
  uid = "UWGAXD",
  pwd = keyring::key_get("oracle_cdr", "UWGAXD"),
  timeout = 10
)



```


---

# Step 1 — Pull Enrollment, Class, Student-Term, and Residency Data

One consolidated query joins:

- `REC_ENROLLMENT_FACT_V` — one row per student per class
- `REC_CLASS_DIM_V` — class metadata (subject, academic group, term)
- `REC_STUDENT_TERM_DIM_V` — student term record (career, load code, academic level)
- `CC_RESIDENCY_DIM_V` — most-recent residency effective on or before the enrollment term

`CTERM_FA_LOAD_CODE` values: **F** = Full load, **N** = No load (audit / non-credit), **T** = 3/4 enrollment, **H** = Half time enrollment, **L** = less than half time.

- `V_STUCRS` from CDR -> UWSA - student rows with YEAR = 202526 (example), TERM = (1 = Fall, 2 = Spring, S = Summer), UNIT = G, CUM_GPA, CLASSIF where A,B,C,D are regular undergrads (freshmen, sophomore, junior, senior) and S = senior over 65. Classif J and O are undergraduate special cases and E is a master's student. This dataset also has RES_STATUS where **R** = Resident, **A** = resident alumni discount, **B** non-resident alumni discount, **V** = non-resident veteran, **L** nonresident but Midwest tuition, **M** = nonresident but Michigan reciprocity, **S** non-resident but Minnesota reciprocity, **W** non-resident but Wisconsin return discount, **E** = nonresident but midwest exchange program, **N** = nonresident general. Importantly, this CDR data can be joined with RDS data where EMPLID variables = CAMPUS_STUDENT_ID.

```{r pull-enrollment}

all_terms_sql <- paste(sprintf("'%s'", ALL_TERMS), collapse = ", ")

stud_enroll_full <- dbGetQuery(
  con_rds,
  sprintf(
    "
    WITH enroll AS (
      SELECT
        e.CTERM_SID,
        e.CLASS_SID,
        e.ENRL_EMPLID,
        e.ENRL_ACAD_CAREER,
        e.ENRL_UNITS_TAKEN,
        e.ENRL_UNITS_BILLING,
        e.ENRL_STATUS_SDESC,
        c.CLASS_TERM_CD,
        c.CLASS_TERM_SDESC,
        c.CLASS_TERM_ACAD_YEAR,
        c.CLASS_ACAD_GROUP_LDESC,   -- e.g. 'Undergraduate', 'Graduate'
        c.CLASS_ACAD_ORG_LDESC,
        c.CLASS_SUBJECT_LDESC
      FROM PKS_RDS_USR.REC_ENROLLMENT_FACT_V e
      JOIN PKS_RDS_USR.REC_CLASS_DIM_V c
        ON c.CLASS_SID = e.CLASS_SID
      WHERE c.CLASS_INSTITUTION_CD = 'UWPKS'
        AND c.CLASS_TERM_CD IN (%s)
    ),
    stud_term AS (
      SELECT
        st.CTERM_SID,
        st.CTERM_EMPLID,
        st.CTERM_TERM_CD,
        st.CTERM_CATEGORY_SDESC,
        st.CTERM_ACAD_YEAR,
        st.CTERM_TERM_SDESC,
        st.CTERM_ACAD_LOAD_SDESC,
        st.CTERM_ACAD_LEVEL_BOT_LDESC,
        st.CTERM_FA_LOAD_CODE         -- F = Full load, N = No load (audit), L = less than full time, T = three quarters, H = half time
      FROM PKS_RDS_USR.REC_STUDENT_TERM_DIM_V st
      WHERE st.CTERM_INSTITUTION_CODE = 'UWPKS'
        AND st.CTERM_TERM_CD IN (%s)
    ),
    res_ranked AS (
      SELECT
        en.ENRL_EMPLID,
        en.CLASS_TERM_CD,
        r.RES_ACAD_CAREER_SDESC,
        r.RES_EFFECTIVE_TERM,
        r.RES_RESIDENCY,
        r.RES_RESIDENCY_LDESC,
        r.RES_TUITION_RES_SDESC,
        r.RES_TUITION_RES_LDESC,
        r.RES_RESIDENCY_DT,
        ROW_NUMBER() OVER (
          PARTITION BY en.ENRL_EMPLID, en.CLASS_TERM_CD
          ORDER BY r.RES_EFFECTIVE_TERM DESC, r.RES_RESIDENCY_DT DESC
        ) AS rn
      FROM enroll en
      LEFT JOIN PKS_RDS_USR.CC_RESIDENCY_DIM_V r
        ON  r.RES_EMPLID      = en.ENRL_EMPLID
        AND r.RES_INSTITUTION = 'UWPKS'
        AND r.RES_EFFECTIVE_TERM <= en.CLASS_TERM_CD
    )
    SELECT
      en.*,
      st.CTERM_CATEGORY_SDESC,
      st.CTERM_ACAD_LOAD_SDESC,
      st.CTERM_ACAD_LEVEL_BOT_LDESC,
      st.CTERM_FA_LOAD_CODE,
      rr.RES_EFFECTIVE_TERM,
      rr.RES_RESIDENCY,
      rr.RES_RESIDENCY_LDESC,
      rr.RES_TUITION_RES_SDESC,
      rr.RES_TUITION_RES_LDESC
    FROM enroll en
    LEFT JOIN stud_term st
      ON  st.CTERM_SID    = en.CTERM_SID
      AND st.CTERM_EMPLID = en.ENRL_EMPLID
      AND st.CTERM_TERM_CD = en.CLASS_TERM_CD
    LEFT JOIN res_ranked rr
      ON  rr.ENRL_EMPLID   = en.ENRL_EMPLID
      AND rr.CLASS_TERM_CD = en.CLASS_TERM_CD
      AND rr.rn = 1
    ",
    all_terms_sql,
    all_terms_sql
  )
)

```

---

# Step 2 — Pull Student Groups (Wisconsin 60+ Identification)

Student group codes identify special populations. We pull all groups for enrolled
students and look for the Senior Citizen / Over-60 code.
Review the values of `STDNT_GROUP` in the output to confirm the exact code used
at UW-Parkside (commonly `'SR60'` or similar; update `WI60_GROUPS` below as needed).


**Student groups really come from Resident Status**

```{r pull-student-groups}
enrolled_emplids_sql <- stud_enroll_full %>%
  distinct(ENRL_EMPLID) %>%
  mutate(ENRL_EMPLID = as.character(ENRL_EMPLID)) %>%
  summarise(ids = paste(sprintf("'%s'", ENRL_EMPLID), collapse = ", ")) %>%
  pull(ids)

stud_groups <- dbGetQuery(
  con_rds,
  sprintf(
    "
    SELECT
      sg.STDNT_GROUP_EMPLID,
      sg.STDNT_GROUP,
      sg.STDNT_GROUP_LDESC
    FROM PKS_RDS_USR.REC_STDNT_GROUPS_DIM_V sg
    WHERE sg.STDNT_GROUP_INSTITUTION = 'UWPKS'
      AND sg.STDNT_GROUP_EMPLID IN (%s)
    ",
    enrolled_emplids_sql
  )
)

# Review unique group codes to confirm the correct 60+ code
stud_groups %>%
  distinct(STDNT_GROUP, STDNT_GROUP_LDESC) %>%
  arrange(STDNT_GROUP)
```

```{r define-wi60-group}
# ── UPDATE THIS VECTOR after reviewing the group codes above ─────────────────
# Common values: 'SR60', 'S', 'OA60' — verify against your stud_groups output
WI60_GROUPS <- c('SR60')   # <-- confirm / update as needed

wi60_emplids <- stud_groups %>%
  filter(STDNT_GROUP %in% WI60_GROUPS) %>%
  distinct(STDNT_GROUP_EMPLID) %>%
  pull(STDNT_GROUP_EMPLID)
```


```{r close_connection}
# Close the connection once all data is retrieved.
dbDisconnect(con_rds)
dbDisconnect(con_cdr)

```


---

# Step 3 — Classify Students and Recode Residency

## 3a — Apply Student Career Groups

Classification priority:

1. **WI 60+ Audit** — student is in the 60+ group, has Wisconsin residency, and
   enrolled as no-load (`CTERM_FA_LOAD_CODE == 'N'`)
2. **UFLEX** — `ENRL_ACAD_CAREER == 'UFLEX'` OR term code does NOT start with `'1'`
3. **GRAD** — `ENRL_ACAD_CAREER == 'GRAD'` (and traditional term)
4. **UGRD** — `ENRL_ACAD_CAREER == 'UGRD'` (and traditional term)
5. **Other** — anything else (review and reclassify as needed)

```{r classify-students}
stud_classified <- stud_enroll_full %>%
  mutate(
    # Wisconsin resident flag (adjust RES_RESIDENCY code to match your data)
    is_wi_resident = RES_RESIDENCY %in% c('RES', 'WIS', 'WIRES'),

    # No-load (audit) flag
    is_no_load = CTERM_FA_LOAD_CODE == 'N',

    # UFlex flag: career code OR term prefix != '1'
    is_uflex = ENRL_ACAD_CAREER == 'UFLEX' |
               !str_starts(as.character(CLASS_TERM_CD), '1'),

    CAREER_GROUP = case_when(
      # 1. Wisconsin 60+ auditors — tracked separately, excluded from billing
      ENRL_EMPLID %in% wi60_emplids & is_wi_resident & is_no_load ~ 'WI 60+ Audit',

      # 2. UFlex
      is_uflex                                                      ~ 'UFLEX',

      # 3. Graduate
      ENRL_ACAD_CAREER == 'GRAD'                                    ~ 'GRAD',

      # 4. Undergraduate
      ENRL_ACAD_CAREER == 'UGRD'                                    ~ 'UGRD',

      TRUE                                                          ~ 'Other'
    )
  )

# Spot-check career group distribution
stud_classified %>%
  count(CAREER_GROUP, ENRL_ACAD_CAREER, sort = TRUE)
```

## 3b — Graduate Students: Keep Graduate-Level Courses Only

Graduate students sometimes enroll in undergraduate courses. For tuition
projection purposes, only credit hours from **graduate-level classes**
(`CLASS_ACAD_GROUP_LDESC` = 'Graduate' or similar) are counted toward the
GRAD bucket. Review the values first.

```{r grad-course-filter}
# Review academic group values
stud_classified %>%
  filter(CAREER_GROUP == 'GRAD') %>%
  count(CLASS_ACAD_GROUP_LDESC, sort = TRUE)
```

```{r apply-grad-filter}
# ── UPDATE the grad_group_values vector to match your data ───────────────────
grad_group_values <- c('Graduate')   # <-- confirm against output above

stud_for_billing <- stud_classified %>%
  filter(
    # For GRAD: only graduate-level courses
    !(CAREER_GROUP == 'GRAD' & !CLASS_ACAD_GROUP_LDESC %in% grad_group_values),
    # Exclude WI 60+ auditors from the billing summary
    CAREER_GROUP != 'WI 60+ Audit'
  )
```

## 3c — Recode Midwest Exchange to Midwest Tuition Non-Resident

UW-Parkside no longer participates in the Midwest Student Exchange. Any student
whose tuition residency description contains "Midwest" and "Exchange" (or
"Compact") is recoded to **Midwest Tuition Non-Resident**.

```{r recode-midwest}
# Review current Midwest-related residency labels before recoding
stud_for_billing %>%
  filter(str_detect(RES_TUITION_RES_LDESC, regex("midwest", ignore_case = TRUE))) %>%
  count(RES_TUITION_RES_LDESC, sort = TRUE)
```

```{r apply-midwest-recode}
stud_for_billing <- stud_for_billing %>%
  mutate(
    RES_TUITION_RES_LDESC = case_when(
      str_detect(
        RES_TUITION_RES_LDESC,
        regex("midwest.*(exchange|compact)|exchange.*(midwest|compact)", ignore_case = TRUE)
      ) ~ "Midwest Tuition Non-Resident",
      TRUE ~ RES_TUITION_RES_LDESC
    )
  )

# Confirm recoding
stud_for_billing %>%
  filter(str_detect(RES_TUITION_RES_LDESC, regex("midwest", ignore_case = TRUE))) %>%
  count(RES_TUITION_RES_LDESC, sort = TRUE)
```

---

# Step 4 — Credit Hours Summary by Term, Career Group, Residency, and Subject

```{r credit-hour-summary}
credit_hours_summary <- stud_for_billing %>%
  group_by(
    CLASS_TERM_ACAD_YEAR,
    CLASS_TERM_SDESC,
    CAREER_GROUP,
    RES_TUITION_RES_LDESC,
    CLASS_SUBJECT_LDESC
  ) %>%
  summarise(
    n_enrollments  = n(),
    credit_hrs_taken   = sum(ENRL_UNITS_TAKEN,   na.rm = TRUE),
    credit_hrs_billing = sum(ENRL_UNITS_BILLING,  na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(CLASS_TERM_ACAD_YEAR, CLASS_TERM_SDESC, CAREER_GROUP,
          RES_TUITION_RES_LDESC, CLASS_SUBJECT_LDESC)

credit_hours_summary
```

---

# Step 5 — Wisconsin 60+ Audit Summary (Separate, Not Billed)

```{r wi60-summary}
wi60_summary <- stud_classified %>%
  filter(CAREER_GROUP == 'WI 60+ Audit') %>%
  group_by(
    CLASS_TERM_ACAD_YEAR,
    CLASS_TERM_SDESC,
    CLASS_SUBJECT_LDESC
  ) %>%
  summarise(
    n_enrollments      = n(),
    credit_hrs_taken   = sum(ENRL_UNITS_TAKEN,  na.rm = TRUE),
    credit_hrs_billing = sum(ENRL_UNITS_BILLING, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(CLASS_TERM_ACAD_YEAR, CLASS_TERM_SDESC, CLASS_SUBJECT_LDESC)

wi60_summary
```

---

# Data Quality Checks

```{r quality-checks}
# Missingness
stud_enroll_full %>%
  summarise(
    n_rows              = n(),
    missing_career      = sum(is.na(ENRL_ACAD_CAREER)),
    missing_term        = sum(is.na(CLASS_TERM_CD)),
    missing_fa_load     = sum(is.na(CTERM_FA_LOAD_CODE)),
    missing_residency   = sum(is.na(RES_RESIDENCY)),
    missing_tuition_res = sum(is.na(RES_TUITION_RES_LDESC)),
    missing_subject     = sum(is.na(CLASS_SUBJECT_LDESC))
  )

# Career group counts
stud_classified %>%
  count(CAREER_GROUP, sort = TRUE)

# Residency distribution in billing data
stud_for_billing %>%
  count(CAREER_GROUP, RES_TUITION_RES_LDESC, sort = TRUE)
```
