---
title: "Credit Hours by Program — Streamlined"
author: "Tony Daza"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(DBI)
library(dplyr)
library(stringr)
library(odbc)
library(keyring)
library(tidyr)
library(ggplot2)
library(knitr)
library(kableExtra)
```

# Overview

This document calculates credit hours by program (subject) for tuition revenue
projections. Students are grouped into the following **career categories**:

| Category | Description |
|---|---|
| **UGRD** | Traditional undergraduate (resident or non-resident) |
| **GRAD** | Graduate students — graduate-level courses only |
| **UFLEX** | University Flex (term code prefix 5 or 6, or `ENRL_ACAD_CAREER == 'UFLEX'`) |
| **WI 65+ Audit** | WI-resident students with `CLASSIF` S or T, enrolled no-load — tracked separately, excluded from billing |

Residency uses CDR `RES_STATUS` codes for traditional students (authoritative
single-character codes). Students with `RES_STATUS = 'E'` (Midwest Exchange)
are recoded to **Midwest Tuition Non-Resident** since UW-Parkside no longer
participates in that program. UFlex students are not in CDR; their residency
falls back to the RDS `RES_TUITION_RES_LDESC` field.

**CDR `RES_STATUS` key:**

| Code | Meaning |
|---|---|
| R | Resident |
| A | Resident — alumni discount |
| B | Non-Resident — alumni discount |
| V | Non-Resident — veteran |
| L | Non-Resident — Midwest tuition |
| M | Non-Resident — Michigan reciprocity |
| S | Non-Resident — Minnesota reciprocity |
| W | Non-Resident — Wisconsin return discount |
| E | Non-Resident — Midwest Exchange *(recoded → Midwest Tuition Non-Resident)* |
| N | Non-Resident — general |

**CDR `CLASSIF` key (from `V_STUCRS`):**

| Code | Meaning |
|---|---|
| A | Freshman |
| B | Sophomore |
| C | Junior |
| D | Senior |
| S | Senior over 65 |
| T | Special classification (potentially 60+; verify) |
| J | Undergraduate special |
| O | Undergraduate special |
| E | Master's student |

---

# Term Reference

```{r term-reference}
# ── Traditional terms (CLASS_TERM_CD begins with 1) ──────────────────────────
# 1257 = Summer 2024–2025
# 1261 = Fall   2025–2026
# 1263 = Winter 2025–2026
# 1265 = Spring 2025–2026
# 1267 = Summer 2025–2026

# ── UFlex terms (CLASS_TERM_CD begins with 5 or 6) ───────────────────────────
# 6506 = Jun–Aug 2025    6507 = Jul–Sep 2025    6508 = Aug–Oct 2025
# 6509 = Sep–Nov 2025    6510 = Oct–Dec 2025    6511 = Nov–Jan 2026
# 6512 = Dec–Feb 2026    6601 = Jan–Mar 2026    6602 = Feb–Apr 2026
# 6603 = Mar–May 2026

TRADITIONAL_TERMS <- c('1257', '1261', '1263', '1265', '1267')
UFLEX_TERMS       <- c('6506', '6507', '6508', '6509', '6510',
                        '6511', '6512', '6601', '6602', '6603')
ALL_TERMS         <- c(TRADITIONAL_TERMS, UFLEX_TERMS)

# ── CDR term mapping (traditional terms only; UFlex not in CDR) ───────────────
# CDR YEAR = 6-digit academic year (e.g., 202526 = AY 2025-2026)
# CDR TERM: 1 = Fall, 2 = Spring, S = Summer
# Note: verify Winter (1263) and Summer (1267) year assignments with CDR team
cdr_term_map <- tribble(
  ~CLASS_TERM_CD, ~CDR_YEAR,  ~CDR_TERM,
  '1257',         '202425',   'S',   # Summer 2024-2025
  '1261',         '202526',   '1',   # Fall   2025-2026
  '1263',         '202526',   'W',   # Winter 2025-2026 (verify if CDR has winter)
  '1265',         '202526',   '2',   # Spring 2025-2026
  '1267',         '202526',   'S'    # Summer 2025-2026 (verify year assignment)
)
```

---

# Connect to Databases

```{r connect}
con_rds <- dbConnect(
  odbc::odbc(),
  dsn     = "Oracle RDS",
  uid     = "DAZA",
  pwd     = keyring::key_get("oracle_rds", "DAZA"),
  timeout = 10
)

con_cdr <- dbConnect(
  odbc::odbc(),
  dsn     = "Oracle CDR",
  uid     = "UWGAXD",
  pwd     = keyring::key_get("oracle_cdr", "UWGAXD"),
  timeout = 10
)
```

---

# Step 1 — Pull Enrollment, Class, Student-Term, and Residency from RDS

One consolidated query joins:

- `REC_ENROLLMENT_FACT_V` — one row per student per class
- `REC_CLASS_DIM_V` — class metadata (subject, academic group, term)
- `REC_STUDENT_TERM_DIM_V` — student term record (career, load code, academic level)
- `CC_RESIDENCY_DIM_V` — most-recent residency effective on or before the enrollment term

`CTERM_FA_LOAD_CODE`: **F** = Full, **T** = Three-quarter, **H** = Half,
**L** = Less than half, **N** = No load (audit / non-credit).

```{r pull-rds}
all_terms_sql <- paste(sprintf("'%s'", ALL_TERMS), collapse = ", ")

stud_enroll_full <- dbGetQuery(
  con_rds,
  sprintf(
    "
    WITH enroll AS (
      SELECT
        e.CTERM_SID,
        e.CLASS_SID,
        e.ENRL_EMPLID,
        e.ENRL_ACAD_CAREER,
        e.ENRL_UNITS_TAKEN,
        e.ENRL_UNITS_BILLING,
        e.ENRL_STATUS_SDESC,
        c.CLASS_TERM_CD,
        c.CLASS_TERM_SDESC,
        c.CLASS_TERM_ACAD_YEAR,
        c.CLASS_ACAD_GROUP_LDESC,   -- e.g. 'Undergraduate', 'Graduate'
        c.CLASS_ACAD_ORG_LDESC,
        c.CLASS_SUBJECT_LDESC
      FROM PKS_RDS_USR.REC_ENROLLMENT_FACT_V e
      JOIN PKS_RDS_USR.REC_CLASS_DIM_V c
        ON c.CLASS_SID = e.CLASS_SID
      WHERE c.CLASS_INSTITUTION_CD = 'UWPKS'
        AND c.CLASS_TERM_CD IN (%s)
    ),
    stud_term AS (
      SELECT
        st.CTERM_SID,
        st.CTERM_EMPLID,
        st.CTERM_TERM_CD,
        st.CTERM_CATEGORY_SDESC,
        st.CTERM_ACAD_YEAR,
        st.CTERM_TERM_SDESC,
        st.CTERM_ACAD_LOAD_SDESC,
        st.CTERM_ACAD_LEVEL_BOT_LDESC,
        st.CTERM_FA_LOAD_CODE         -- F = Full load, N = No load (audit), L = less than full time, T = three quarters, H = half time
      FROM PKS_RDS_USR.REC_STUDENT_TERM_DIM_V st
      WHERE st.CTERM_INSTITUTION_CODE = 'UWPKS'
        AND st.CTERM_TERM_CD IN (%s)
    ),
    res_ranked AS (
      SELECT
        en.ENRL_EMPLID,
        en.CLASS_TERM_CD,
        r.RES_ACAD_CAREER_SDESC,
        r.RES_EFFECTIVE_TERM,
        r.RES_RESIDENCY,
        r.RES_RESIDENCY_LDESC,
        r.RES_TUITION_RES_SDESC,
        r.RES_TUITION_RES_LDESC,
        r.RES_RESIDENCY_DT,
        ROW_NUMBER() OVER (
          PARTITION BY en.ENRL_EMPLID, en.CLASS_TERM_CD
          ORDER BY r.RES_EFFECTIVE_TERM DESC, r.RES_RESIDENCY_DT DESC
        ) AS rn
      FROM enroll en
      LEFT JOIN PKS_RDS_USR.CC_RESIDENCY_DIM_V r
        ON  r.RES_EMPLID      = en.ENRL_EMPLID
        AND r.RES_INSTITUTION = 'UWPKS'
        AND r.RES_EFFECTIVE_TERM <= en.CLASS_TERM_CD
    )
    SELECT
      en.*,
      st.CTERM_CATEGORY_SDESC,
      st.CTERM_ACAD_LOAD_SDESC,
      st.CTERM_ACAD_LEVEL_BOT_LDESC,
      st.CTERM_FA_LOAD_CODE,
      rr.RES_EFFECTIVE_TERM,
      rr.RES_RESIDENCY,
      rr.RES_RESIDENCY_LDESC,
      rr.RES_TUITION_RES_SDESC,
      rr.RES_TUITION_RES_LDESC
    FROM enroll en
    LEFT JOIN stud_term st
      ON  st.CTERM_SID    = en.CTERM_SID
      AND st.CTERM_EMPLID = en.ENRL_EMPLID
      AND st.CTERM_TERM_CD = en.CLASS_TERM_CD
    LEFT JOIN res_ranked rr
      ON  rr.ENRL_EMPLID   = en.ENRL_EMPLID
      AND rr.CLASS_TERM_CD = en.CLASS_TERM_CD
      AND rr.rn = 1
    ",
    all_terms_sql,
    all_terms_sql
  )
)
```

---

# Step 2 — Pull Student Classification and Residency from CDR (`V_STUCRS`)

CDR `V_STUCRS` (UWSA) provides authoritative `CLASSIF` and `RES_STATUS` for
traditional-term students. UFlex students are **not** in CDR.

Join key: CDR `CAMPUS_STUDENT_ID` = RDS `ENRL_EMPLID`.

```{r pull-cdr}
# Build enrolled EMPLID list for traditional terms only
trad_emplids <- stud_enroll_full %>%
  filter(str_starts(as.character(CLASS_TERM_CD), "1")) %>%
  distinct(ENRL_EMPLID) %>%
  mutate(ENRL_EMPLID = as.character(ENRL_EMPLID)) %>%
  pull(ENRL_EMPLID)

trad_emplids_sql <- paste(sprintf("'%s'", trad_emplids), collapse = ", ")

# Build (YEAR, TERM) pairs from the mapping table for the WHERE clause
cdr_year_term_pairs <- cdr_term_map %>%
  mutate(pair = sprintf("(YEAR = '%s' AND TERM = '%s')", CDR_YEAR, CDR_TERM)) %>%
  pull(pair) %>%
  paste(collapse = " OR ")

emplid_chunks <- split(trad_emplids, ceiling(seq_along(trad_emplids) / 1000))

stud_cdr <- purrr::map_dfr(emplid_chunks, function(chunk) {
  chunk_sql <- paste(sprintf("'%s'", chunk), collapse = ", ")
  
  dbGetQuery(
    con_cdr,
    sprintf(
      "
      SELECT
        sc.CAMPUS_STUDENT_ID,
        sc.YEAR,
        sc.TERM,
        sc.CLASSIF,
        sc.RES_STATUS
      FROM UWSA.V_STUCRS sc
      WHERE sc.UNIT = 'G'
        AND sc.CAMPUS_STUDENT_ID IN (%s)
        AND (%s)
      ",
      chunk_sql,
      cdr_year_term_pairs
    )
  )
})

# Join CDR term identifiers back to RDS CLASS_TERM_CD for merging
stud_cdr <- stud_cdr %>%
  mutate(YEAR = as.character(YEAR), TERM = as.character(TERM)) %>%
  left_join(
    cdr_term_map %>% mutate(YEAR = CDR_YEAR, TERM = CDR_TERM),
    by = c("YEAR", "TERM")
  ) %>%
  select(CAMPUS_STUDENT_ID, CLASS_TERM_CD, CLASSIF, RES_STATUS)

# Spot-check CDR values
cat("CLASSIF distribution:\n")
stud_cdr %>% count(CLASSIF, sort = TRUE) %>% print()
cat("\nRES_STATUS distribution:\n")
stud_cdr %>% count(RES_STATUS, sort = TRUE) %>% print()
```

```{r disconnect}
# Close connections — all data is now in R
dbDisconnect(con_rds)
dbDisconnect(con_cdr)
```

---

# Step 3 — Join RDS and CDR Data

```{r join-rds-cdr}
stud_combined <- stud_enroll_full %>%
  left_join(
    stud_cdr,
    by = c("ENRL_EMPLID" = "CAMPUS_STUDENT_ID", "CLASS_TERM_CD")
  )

# CDR fields will be NA for UFlex students (expected — they are not in CDR)
stud_combined %>%
  mutate(is_uflex = !str_starts(as.character(CLASS_TERM_CD), "1")) %>%
  group_by(is_uflex) %>%
  summarise(
    n_rows          = n(),
    missing_classif = sum(is.na(CLASSIF)),
    missing_res_status = sum(is.na(RES_STATUS))
  )

stud_combined %>% count(ENRL_ACAD_CAREER, sprt = TRUE)
```

---

# Step 4 — Classify Students

## 4a — Career Groups

Classification priority (applied in order):

1. **WI 65+ Audit** — `CLASSIF %in% c('S', 'T')` AND `RES_STATUS == 'R'` AND
   `CTERM_FA_LOAD_CODE == 'N'`
2. **UFLEX** — `ENRL_ACAD_CAREER == 'UFLEX'` OR term code does not start with `'1'`
3. **GRAD** — `ENRL_ACAD_CAREER == 'GRAD'`
4. **UGRD** — `ENRL_ACAD_CAREER == 'UGRD'`
5. **Other** — review and reclassify as needed

```{r classify-career}
stud_classified <- stud_combined %>%
  mutate(
    is_senior_65  = CLASSIF %in% c('S', 'T'),
    is_wi_res_cdr = RES_STATUS == 'R',
    is_no_load    = CTERM_FA_LOAD_CODE == 'N',
    is_uflex      = ENRL_ACAD_CAREER == 'UFLX',

    CAREER_GROUP = case_when(
      is_senior_65 & is_wi_res_cdr & is_no_load ~ 'WI 65+ Audit',
      is_uflex                                  ~ 'UFLX',
      ENRL_ACAD_CAREER == 'GRAD'                ~ 'GRAD',
      ENRL_ACAD_CAREER == 'UGRD'                ~ 'UGRD',
      TRUE                                      ~ 'Other'
    )
  )

stud_classified %>%
  count(CAREER_GROUP, ENRL_ACAD_CAREER, sort = TRUE)
```

## 4b — Graduate Students: Keep Graduate-Level Courses Only

Grad students sometimes take undergraduate courses. Only credit hours from
graduate-level classes count toward the GRAD billing bucket.

```{r grad-filter}
# Review academic group labels for GRAD students before filtering
stud_classified %>%
  filter(CAREER_GROUP == 'GRAD') %>%
  count(CLASS_ACAD_GROUP_LDESC, sort = TRUE)
```

```{r apply-grad-filter}
# Confirm 'Graduate' is the correct label in your data and update if needed
GRAD_COURSE_GROUPS <- c('Graduate')

stud_for_billing <- stud_classified %>%
  filter(
    !(CAREER_GROUP == 'GRAD' & !CLASS_ACAD_GROUP_LDESC %in% GRAD_COURSE_GROUPS),
    CAREER_GROUP != 'WI 65+ Audit'
  )
```

## 4c — Assign Residency Category

For **traditional students**, CDR `RES_STATUS` is the authoritative source.
`RES_STATUS = 'E'` (Midwest Exchange) is recoded to **Midwest Tuition Non-Resident**.

For **UFlex students**, `RES_STATUS` will be `NA` (not in CDR), so we fall
back to the RDS `RES_TUITION_RES_LDESC` field and apply the same Exchange recode
via text matching as a safety net.

```{r assign-residency}


stud_for_billing <- stud_for_billing %>%
  mutate(
    RESIDENCY_CATEGORY = case_when(

      # ── Traditional students: use CDR RES_STATUS ───────────────────────────
      !is_uflex & RES_STATUS == 'R' ~ 'Resident',
      !is_uflex & RES_STATUS == 'A' ~ 'Resident - Alumni Discount',
      !is_uflex & RES_STATUS == 'E' ~ 'Midwest Tuition Non-Resident',   # recoded from Exchange
      !is_uflex & RES_STATUS == 'L' ~ 'Midwest Tuition Non-Resident',
      !is_uflex & RES_STATUS == 'M' ~ 'Non-Resident - Michigan Reciprocity',
      !is_uflex & RES_STATUS == 'S' ~ 'Non-Resident - Minnesota Reciprocity',
      !is_uflex & RES_STATUS == 'W' ~ 'Non-Resident - Wisconsin Return Discount',
      !is_uflex & RES_STATUS == 'V' ~ 'Non-Resident - Veteran',
      !is_uflex & RES_STATUS == 'B' ~ 'Non-Resident - Alumni Discount',
      !is_uflex & RES_STATUS == 'N' ~ 'Non-Resident',
      !is_uflex & !is.na(RES_STATUS) ~ paste0('Non-Resident - Other (', RES_STATUS, ')'),

      # ── UFlex students: fall back to RDS residency description ─────────────
      is_uflex ~ case_when(
        str_detect(
          RES_TUITION_RES_LDESC,
          regex("midwest.*(exchange|compact)|exchange.*(midwest|compact)", ignore_case = TRUE)
        )                            ~ 'Midwest Tuition Non-Resident',
        !is.na(RES_TUITION_RES_LDESC) ~ RES_TUITION_RES_LDESC,
        TRUE                         ~ 'Unknown'
      ),

      TRUE ~ 'Unknown'
    )
  )

getOption("na.print")  # see what it's currently set to
options(na.print = "NA")  # reset to default

# Verify residency category distribution
stud_for_billing %>%
  count(CAREER_GROUP, RESIDENCY_CATEGORY, sort = TRUE) %>%
  as_tibble() %>%
  print(n = 40)

```

---

# Step 5 — Credit Hours Summary

```{r credit-hour-summary}
credit_hours_summary <- stud_for_billing %>%
  group_by(
    CLASS_TERM_ACAD_YEAR,
    CLASS_TERM_SDESC,
    CAREER_GROUP,
    RESIDENCY_CATEGORY,
    CLASS_SUBJECT_LDESC
  ) %>%
  summarise(
    n_enrollments      = n(),
    credit_hrs_taken   = sum(ENRL_UNITS_TAKEN,   na.rm = TRUE),
    credit_hrs_billing = sum(ENRL_UNITS_BILLING,  na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(CLASS_TERM_ACAD_YEAR, CLASS_TERM_SDESC, CAREER_GROUP,
          RESIDENCY_CATEGORY, CLASS_SUBJECT_LDESC)

credit_hours_summary
```

---

# Step 6 — WI 65+ Audit Summary (Separate, Not Billed)

```{r wi65-summary}
wi65_summary <- stud_classified %>%
  filter(CAREER_GROUP == 'WI 65+ Audit') %>%
  group_by(
    CLASS_TERM_ACAD_YEAR,
    CLASS_TERM_SDESC,
    CLASS_SUBJECT_LDESC
  ) %>%
  summarise(
    n_enrollments      = n(),
    credit_hrs_taken   = sum(ENRL_UNITS_TAKEN,  na.rm = TRUE),
    credit_hrs_billing = sum(ENRL_UNITS_BILLING, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(CLASS_TERM_ACAD_YEAR, CLASS_TERM_SDESC, CLASS_SUBJECT_LDESC)

wi65_summary
```

---

# Data Quality Checks

```{r quality-checks}
# Overall missingness
stud_enroll_full %>%
  summarise(
    n_rows              = n(),
    missing_career      = sum(is.na(ENRL_ACAD_CAREER)),
    missing_fa_load     = sum(is.na(CTERM_FA_LOAD_CODE)),
    missing_residency   = sum(is.na(RES_RESIDENCY)),
    missing_tuition_res = sum(is.na(RES_TUITION_RES_LDESC)),
    missing_subject     = sum(is.na(CLASS_SUBJECT_LDESC))
  )

# CDR join coverage (expect NA only for UFlex)
stud_classified %>%
  group_by(CAREER_GROUP) %>%
  summarise(
    n              = n(),
    missing_classif    = sum(is.na(CLASSIF)),
    missing_res_status = sum(is.na(RES_STATUS))
  )

# Career group counts
stud_classified %>%
  count(CAREER_GROUP, sort = TRUE)

# Any 'Other' career group rows to investigate
stud_classified %>%
  filter(CAREER_GROUP == 'Other') %>%
  count(ENRL_ACAD_CAREER, CLASS_TERM_CD, sort = TRUE)

# Residency distribution in billing data
stud_for_billing %>%
  count(CAREER_GROUP, RESIDENCY_CATEGORY, sort = TRUE)
```
