---
title: "Credit Hours by Program — Streamlined"
author: "Tony Daza"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(DBI)
library(dplyr)
library(stringr)
library(odbc)
library(keyring)
library(tidyr)
library(ggplot2)
library(knitr)
library(kableExtra)
```

# Overview

This document calculates credit hours by program (subject) for tuition revenue
projections. Students are grouped into the following **career categories**:

| Category | Description |
|---|---|
| **UGRD** | Traditional undergraduate (resident or non-resident) | `CLASSIF` = A, B, C, D and maybe J, N and O
| **GRAD** | Graduate students — graduate-level courses only | `CLASSIF` = E and maybe K, G, H
| **UFLX** | University Flex (term code prefix 5 or 6, or `ENRL_ACAD_CAREER == 'UFLX'`) |
| **WI 65+ Audit** | WI-resident students with `CLASSIF` S or T, enrolled no-load — tracked separately, excluded from billing |

Residency uses CDR `RES_STATUS` codes for traditional students (authoritative
single-character codes). Students with `RES_STATUS = 'E'` (Midwest Exchange)
are recoded to **Midwest Tuition Non-Resident** since UW-Parkside no longer
participates in that program. UFlex students are not in CDR; their residency
falls back to the RDS `RES_TUITION_RES_LDESC` field.

However we must also examine if a student has a any residency status, even if not for the specific term when grouping then select the most recent `RES_STATUS`. We will have to use `YEAR` as unfortunately there is no effective date. So essentially, if the student observation has `CLASSIF` and `RES_STATUS` info for the `YEAR` then use that, if not, pull forward the most recent observation for that student. 

**CDR `RES_STATUS` key:**

| Code | Meaning |
|---|---|
| R | Resident |
| A | Resident — alumni discount |
| B | Non-Resident — alumni discount |
| V | Non-Resident — veteran |
| L | Non-Resident — Midwest tuition |
| M | Non-Resident — Michigan reciprocity |
| S | Non-Resident — Minnesota reciprocity |
| W | Non-Resident — Wisconsin return discount |
| E | Non-Resident — Midwest Exchange *(recoded → Midwest Tuition Non-Resident)* |
| N | Non-Resident — general |

**CDR `CLASSIF` key (from `V_STUCRS`):**

| Code | Meaning |
|---|---|
| A | Freshman |
| B | Sophomore |
| C | Junior |
| D | Senior |
| S | Senior over 65 |
| T | Special classification (potentially 60+; verify) |
| J | Undergraduate special |
| N | Undergraduate special |
| O | Undergraduate special |
| E | Master's student |

---

# Term Reference

```{r term-reference}
# ── Traditional terms (CLASS_TERM_CD begins with 1) ──────────────────────────
# 1257 = Summer 2024–2025
# 1261 = Fall   2025–2026
# 1263 = Winter 2025–2026
# 1265 = Spring 2025–2026
# 1267 = Summer 2025–2026

# ── UFlex terms (CLASS_TERM_CD begins with 5 or 6) ───────────────────────────
# 6506 = Jun–Aug 2025    6507 = Jul–Sep 2025    6508 = Aug–Oct 2025
# 6509 = Sep–Nov 2025    6510 = Oct–Dec 2025    6511 = Nov–Jan 2026
# 6512 = Dec–Feb 2026    6601 = Jan–Mar 2026    6602 = Feb–Apr 2026
# 6603 = Mar–May 2026

TRADITIONAL_TERMS <- c('1257', '1261', '1263', '1265', '1267')
UFLEX_TERMS       <- c('6506', '6507', '6508', '6509', '6510',
                        '6511', '6512', '6601', '6602', '6603')
ALL_TERMS         <- c(TRADITIONAL_TERMS, UFLEX_TERMS)

# ── CDR term mapping (traditional terms only; UFlex not in CDR) ───────────────
# CDR YEAR = 6-digit academic year (e.g., 202526 = AY 2025-2026)
# CDR TERM: 1 = Fall, 2 = Spring, S = Summer
# Note: verify Winter (1263) and Summer (1267) year assignments with CDR team
cdr_term_map <- tribble(
  ~CLASS_TERM_CD, ~CDR_YEAR,  ~CDR_TERM,
  '1257',         '202425',   'S',   # Summer 2024-2025
  '1261',         '202526',   '1',   # Fall   2025-2026
  '1263',         '202526',   'W',   # Winter 2025-2026 (verify if CDR has winter)
  '1265',         '202526',   '2',   # Spring 2025-2026
  '1267',         '202526',   'S'    # Summer 2025-2026 (verify year assignment)
)
```

---

# Connect to Databases

```{r connect}
con_rds <- dbConnect(
  odbc::odbc(),
  dsn     = "Oracle RDS",
  uid     = "DAZA",
  pwd     = keyring::key_get("oracle_rds", "DAZA"),
  timeout = 10
)

con_cdr <- dbConnect(
  odbc::odbc(),
  dsn     = "Oracle CDR",
  uid     = "UWGAXD",
  pwd     = keyring::key_get("oracle_cdr", "UWGAXD"),
  timeout = 10
)
```

---

# Step 1 — Pull Enrollment, Class, Student-Term, and Residency from RDS

One consolidated query joins:

- `REC_ENROLLMENT_FACT_V` — one row per student per class
- `REC_CLASS_DIM_V` — class metadata (subject, academic group, term)
- `REC_STUDENT_TERM_DIM_V` — student term record (career, load code, academic level)
- `CC_RESIDENCY_DIM_V` — most-recent residency effective on or before the enrollment term

`CTERM_FA_LOAD_CODE`: **F** = Full, **T** = Three-quarter, **H** = Half,
**L** = Less than half, **N** = No load (audit / non-credit).

```{r pull-rds}
all_terms_sql <- paste(sprintf("'%s'", ALL_TERMS), collapse = ", ")

stud_enroll_full <- dbGetQuery(
  con_rds,
  sprintf(
    "
    WITH enroll AS (
      SELECT
        e.CTERM_SID,
        e.CLASS_SID,
        e.ENRL_EMPLID,
        e.ENRL_ACAD_CAREER,
        e.ENRL_UNITS_TAKEN,
        e.ENRL_UNITS_BILLING,
        e.ENRL_STATUS_SDESC,
        c.CLASS_TERM_CD,
        c.CLASS_TERM_SDESC,
        c.CLASS_TERM_ACAD_YEAR,
        c.CLASS_ACAD_GROUP_LDESC,   -- e.g. 'Undergraduate', 'Graduate'
        c.CLASS_ACAD_ORG_LDESC,
        c.CLASS_SUBJECT_LDESC
      FROM PKS_RDS_USR.REC_ENROLLMENT_FACT_V e
      JOIN PKS_RDS_USR.REC_CLASS_DIM_V c
        ON c.CLASS_SID = e.CLASS_SID
      WHERE c.CLASS_INSTITUTION_CD = 'UWPKS'
        AND c.CLASS_TERM_CD IN (%s)
    ),
    stud_term AS (
      SELECT
        st.CTERM_SID,
        st.CTERM_EMPLID,
        st.CTERM_TERM_CD,
        st.CTERM_CATEGORY_SDESC,
        st.CTERM_ACAD_YEAR,
        st.CTERM_TERM_SDESC,
        st.CTERM_ACAD_LOAD_SDESC,
        st.CTERM_ACAD_LEVEL_BOT_LDESC,
        st.CTERM_FA_LOAD_CODE         -- F = Full load, N = No load (audit), L = less than full time, T = three quarters, H = half time
      FROM PKS_RDS_USR.REC_STUDENT_TERM_DIM_V st
      WHERE st.CTERM_INSTITUTION_CODE = 'UWPKS'
        AND st.CTERM_TERM_CD IN (%s)
    ),
    res_ranked AS (
      SELECT
        en.ENRL_EMPLID,
        en.CLASS_TERM_CD,
        r.RES_ACAD_CAREER_SDESC,
        r.RES_EFFECTIVE_TERM,
        r.RES_RESIDENCY,
        r.RES_RESIDENCY_LDESC,
        r.RES_TUITION_RES_SDESC,
        r.RES_TUITION_RES_LDESC,
        r.RES_RESIDENCY_DT,
        ROW_NUMBER() OVER (
          PARTITION BY en.ENRL_EMPLID, en.CLASS_TERM_CD
          ORDER BY r.RES_EFFECTIVE_TERM DESC, r.RES_RESIDENCY_DT DESC
        ) AS rn
      FROM enroll en
      LEFT JOIN PKS_RDS_USR.CC_RESIDENCY_DIM_V r
        ON  r.RES_EMPLID      = en.ENRL_EMPLID
        AND r.RES_INSTITUTION = 'UWPKS'
        AND r.RES_EFFECTIVE_TERM <= en.CLASS_TERM_CD
    )
    SELECT
      en.*,
      st.CTERM_CATEGORY_SDESC,
      st.CTERM_ACAD_LOAD_SDESC,
      st.CTERM_ACAD_LEVEL_BOT_LDESC,
      st.CTERM_FA_LOAD_CODE,
      rr.RES_EFFECTIVE_TERM,
      rr.RES_RESIDENCY,
      rr.RES_RESIDENCY_LDESC,
      rr.RES_TUITION_RES_SDESC,
      rr.RES_TUITION_RES_LDESC
    FROM enroll en
    LEFT JOIN stud_term st
      ON  st.CTERM_SID    = en.CTERM_SID
      AND st.CTERM_EMPLID = en.ENRL_EMPLID
      AND st.CTERM_TERM_CD = en.CLASS_TERM_CD
    LEFT JOIN res_ranked rr
      ON  rr.ENRL_EMPLID   = en.ENRL_EMPLID
      AND rr.CLASS_TERM_CD = en.CLASS_TERM_CD
      AND rr.rn = 1
    ",
    all_terms_sql,
    all_terms_sql
  )
)
```

---

# Step 2 — Pull Student Classification and Residency from CDR (`V_STUCRS`)

CDR `V_STUCRS` (UWSA) provides authoritative `CLASSIF` and `RES_STATUS` for
traditional-term students. UFlex students are **not** in CDR.

Join key: CDR `CAMPUS_STUDENT_ID` = RDS `ENRL_EMPLID`.

Two queries are issued per student chunk:

1. **In-scope terms** — the specific `(YEAR, TERM)` pairs matching the
   traditional terms being analysed.
2. **Full history** — all `(YEAR, TERM)` records for the same students, used
   in Step 2b to fill forward any `NA` `CLASSIF` or `RES_STATUS` values from
   the student's most recent prior term.

```{r pull-cdr}
# Build enrolled EMPLID list for traditional terms only
trad_emplids <- stud_enroll_full %>%
  filter(str_starts(as.character(CLASS_TERM_CD), "1")) %>%
  distinct(ENRL_EMPLID) %>%
  mutate(ENRL_EMPLID = as.character(ENRL_EMPLID)) %>%
  pull(ENRL_EMPLID)

trad_emplids_sql <- paste(sprintf("'%s'", trad_emplids), collapse = ", ")

# Build (YEAR, TERM) pairs from the mapping table for the WHERE clause
cdr_year_term_pairs <- cdr_term_map %>%
  mutate(pair = sprintf("(YEAR = '%s' AND TERM = '%s')", CDR_YEAR, CDR_TERM)) %>%
  pull(pair) %>%
  paste(collapse = " OR ")

emplid_chunks <- split(trad_emplids, ceiling(seq_along(trad_emplids) / 1000))

stud_cdr <- purrr::map_dfr(emplid_chunks, function(chunk) {
  chunk_sql <- paste(sprintf("'%s'", chunk), collapse = ", ")
  
  dbGetQuery(
    con_cdr,
    sprintf(
      "
      SELECT
        sc.CAMPUS_STUDENT_ID,
        sc.YEAR,
        sc.TERM,
        sc.CLASSIF,
        sc.RES_STATUS
      FROM UWSA.V_STUCRS sc
      WHERE sc.UNIT = 'G'
        AND sc.CAMPUS_STUDENT_ID IN (%s)
        AND (%s)
      ",
      chunk_sql,
      cdr_year_term_pairs
    )
  )
})

# Join CDR term identifiers back to RDS CLASS_TERM_CD for merging
stud_cdr <- stud_cdr %>%
  mutate(YEAR = as.character(YEAR), TERM = as.character(TERM)) %>%
  left_join(
    cdr_term_map %>% mutate(YEAR = CDR_YEAR, TERM = CDR_TERM),
    by = c("YEAR", "TERM")
  ) %>%
  select(CAMPUS_STUDENT_ID, CLASS_TERM_CD, CLASSIF, RES_STATUS) %>%
  # V_STUCRS is course-grain; deduplicate to one row per student x term
  # so the subsequent left join does not create row-multiplying duplicates
  distinct()

# Spot-check CDR values for in-scope terms (pre fill-forward baseline)
cat("CLASSIF distribution (in-scope terms, before fill-forward):\n")
stud_cdr %>% count(CLASSIF, sort = TRUE) %>% print()
cat("\nRES_STATUS distribution (in-scope terms, before fill-forward):\n")
stud_cdr %>% count(RES_STATUS, sort = TRUE) %>% print()

# ── Secondary pull: all CDR records for these students (fill-forward) ────────
# Fetches ALL years/terms so we can backfill CLASSIF and RES_STATUS when a
# student has no record — or a NA value — for a specific in-scope term.
stud_cdr_history <- purrr::map_dfr(emplid_chunks, function(chunk) {
  chunk_sql <- paste(sprintf("'%s'", chunk), collapse = ", ")

  dbGetQuery(
    con_cdr,
    sprintf(
      "
      SELECT
        sc.CAMPUS_STUDENT_ID,
        sc.YEAR,
        sc.TERM,
        sc.CLASSIF,
        sc.RES_STATUS
      FROM UWSA.V_STUCRS sc
      WHERE sc.UNIT = 'G'
        AND sc.CAMPUS_STUDENT_ID IN (%s)
      ",
      chunk_sql
    )
  )
}) %>%
  mutate(YEAR = as.character(YEAR), TERM = as.character(TERM))
```

```{r disconnect}
# Close connections — all data is now in R
dbDisconnect(con_rds)
dbDisconnect(con_cdr)
```

---

# Step 2b — Fill Forward Missing CLASSIF and RES_STATUS from Prior Terms

For traditional students, the CDR `V_STUCRS` record for a specific term may
be absent or have a `NA` value for `CLASSIF` or `RES_STATUS`.  The rules are:

- **CLASSIF**: if `NA` for the in-scope term, use the value from the student's
  most recent *prior* CDR term that has a non-`NA` `CLASSIF`.
- **RES_STATUS**: if `NA` for the in-scope term, use the value from the
  student's most recent *prior* CDR term that has a non-`NA` `RES_STATUS`.

Ordering uses a composite `YEAR_TERM_SORT` key:
`YEAR × 10 + term-position`, where term positions are
Fall (1) < Winter (2) < Spring (3) < Summer (4).

```{r fill-forward-cdr}
# ── Helper: build YEAR_TERM_SORT from character YEAR and TERM columns ────────
make_year_term_sort <- function(year_chr, term_chr) {
  term_num <- case_when(
    term_chr == '1' ~ 1L,   # Fall
    term_chr == 'W' ~ 2L,   # Winter
    term_chr == '2' ~ 3L,   # Spring
    term_chr == 'S' ~ 4L,   # Summer
    TRUE            ~ 9L
  )
  as.integer(year_chr) * 10L + term_num
}

# Sort key for every in-scope term (links CLASS_TERM_CD to a comparable integer)
scope_sort <- cdr_term_map %>%
  mutate(YEAR_TERM_SORT = make_year_term_sort(CDR_YEAR, CDR_TERM)) %>%
  select(CLASS_TERM_CD, YEAR_TERM_SORT)

# Add the sort key to the full CDR history
stud_cdr_history <- stud_cdr_history %>%
  mutate(YEAR_TERM_SORT = make_year_term_sort(YEAR, TERM))

# ── CLASSIF lookup: one non-NA value per student × historical term ───────────
# If multiple records exist for the same student × term, keep the first
# non-NA CLASSIF (CLASSIF is a student-level attribute; multiple rows arise
# because V_STUCRS is course-grain).
classif_lookup <- stud_cdr_history %>%
  filter(!is.na(CLASSIF)) %>%
  distinct(CAMPUS_STUDENT_ID, YEAR_TERM_SORT, CLASSIF) %>%
  group_by(CAMPUS_STUDENT_ID, YEAR_TERM_SORT) %>%
  slice(1L) %>%
  ungroup()

# ── RES_STATUS lookup: one non-NA value per student × historical term ────────
res_status_lookup <- stud_cdr_history %>%
  filter(!is.na(RES_STATUS)) %>%
  distinct(CAMPUS_STUDENT_ID, YEAR_TERM_SORT, RES_STATUS) %>%
  group_by(CAMPUS_STUDENT_ID, YEAR_TERM_SORT) %>%
  slice(1L) %>%
  ungroup()

# Attach the sort key to the in-scope CDR data
stud_cdr <- stud_cdr %>%
  left_join(scope_sort, by = "CLASS_TERM_CD")

# ── Fill-forward CLASSIF ──────────────────────────────────────────────────────
# For each student × term where CLASSIF is NA, cross-join to all prior
# historical CLASSIF records (HIST_SORT < current YEAR_TERM_SORT), then
# keep only the single most-recent prior match.
classif_ff <- stud_cdr %>%
  filter(is.na(CLASSIF)) %>%
  select(CAMPUS_STUDENT_ID, CLASS_TERM_CD, YEAR_TERM_SORT) %>%
  left_join(
    classif_lookup %>%
      rename(HIST_SORT = YEAR_TERM_SORT, CLASSIF_FF = CLASSIF),
    by           = "CAMPUS_STUDENT_ID",
    relationship = "many-to-many"
  ) %>%
  filter(HIST_SORT < YEAR_TERM_SORT) %>%
  group_by(CAMPUS_STUDENT_ID, CLASS_TERM_CD) %>%
  slice_max(HIST_SORT, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(CAMPUS_STUDENT_ID, CLASS_TERM_CD, CLASSIF_FF)

# ── Fill-forward RES_STATUS ───────────────────────────────────────────────────
# Same pattern as CLASSIF: use the most recent prior term with a non-NA value.
res_status_ff <- stud_cdr %>%
  filter(is.na(RES_STATUS)) %>%
  select(CAMPUS_STUDENT_ID, CLASS_TERM_CD, YEAR_TERM_SORT) %>%
  left_join(
    res_status_lookup %>%
      rename(HIST_SORT = YEAR_TERM_SORT, RES_STATUS_FF = RES_STATUS),
    by           = "CAMPUS_STUDENT_ID",
    relationship = "many-to-many"
  ) %>%
  filter(HIST_SORT < YEAR_TERM_SORT) %>%
  group_by(CAMPUS_STUDENT_ID, CLASS_TERM_CD) %>%
  slice_max(HIST_SORT, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(CAMPUS_STUDENT_ID, CLASS_TERM_CD, RES_STATUS_FF)

# ── Merge fill-forward values back into stud_cdr and clean up ────────────────
stud_cdr <- stud_cdr %>%
  left_join(classif_ff,    by = c("CAMPUS_STUDENT_ID", "CLASS_TERM_CD")) %>%
  left_join(res_status_ff, by = c("CAMPUS_STUDENT_ID", "CLASS_TERM_CD")) %>%
  mutate(
    CLASSIF    = coalesce(CLASSIF,    CLASSIF_FF),
    RES_STATUS = coalesce(RES_STATUS, RES_STATUS_FF)
  ) %>%
  select(-YEAR_TERM_SORT, -CLASSIF_FF, -RES_STATUS_FF)

cat("CLASSIF fill-forward:    applied to", nrow(classif_ff),    "student × term rows\n")
cat("RES_STATUS fill-forward: applied to", nrow(res_status_ff), "student × term rows\n")

cat("\nCLASSIF distribution (after fill-forward):\n")
stud_cdr %>% count(CLASSIF, sort = TRUE) %>% print()
cat("\nRES_STATUS distribution (after fill-forward):\n")
stud_cdr %>% count(RES_STATUS, sort = TRUE) %>% print()
```

---

# Step 3 — Join RDS and CDR Data

```{r join-rds-cdr}
stud_combined <- stud_enroll_full %>%
  # Oracle returns EMPLID as numeric; CDR returns CAMPUS_STUDENT_ID as character.
  # Coerce before joining to avoid a type-mismatch silent non-match.
  mutate(ENRL_EMPLID = as.character(ENRL_EMPLID)) %>%
  left_join(
    stud_cdr,
    by = c("ENRL_EMPLID" = "CAMPUS_STUDENT_ID", "CLASS_TERM_CD")
  )

# CDR fields will be NA for UFlex students (expected — they are not in CDR)
stud_combined %>%
  mutate(is_uflex = !str_starts(as.character(CLASS_TERM_CD), "1")) %>%
  group_by(is_uflex) %>%
  summarise(
    n_rows          = n(),
    missing_classif = sum(is.na(CLASSIF)),
    missing_res_status = sum(is.na(RES_STATUS))
  )

stud_combined %>% count(ENRL_ACAD_CAREER, sprt = TRUE)
```

---

# Step 4 — Classify Students

## 4a — Career Groups

Classification priority (applied in order):

1. **WI 65+ Audit** — `CLASSIF %in% c('S', 'T')` AND `RES_STATUS == 'R'` AND
   `CTERM_FA_LOAD_CODE == 'N'`
2. **UFLX** — `ENRL_ACAD_CAREER == 'UFLX'` OR term code does not start with `'1'`
3. **GRAD** — `ENRL_ACAD_CAREER == 'GRAD'`
4. **UGRD** — `ENRL_ACAD_CAREER == 'UGRD'`
5. **Other** — review and reclassify as needed

```{r classify-career}
stud_classified <- stud_combined %>%
  mutate(
    is_senior_65  = CLASSIF %in% c('S', 'T'),
    is_wi_res_cdr = RES_STATUS == 'R',
    is_no_load    = CTERM_FA_LOAD_CODE == 'N',
    # Identify UFlex by BOTH career code ('UFLX') AND term-code prefix (5/6),
    # so UFlex students enrolled under an unexpected career code are still caught.
    is_uflex      = ENRL_ACAD_CAREER == 'UFLX' |
                    !str_starts(as.character(CLASS_TERM_CD), "1"),

    CAREER_GROUP = case_when(
      is_senior_65 & is_wi_res_cdr & is_no_load ~ 'WI 65+ Audit',
      is_uflex                                  ~ 'UFLX',
      ENRL_ACAD_CAREER == 'GRAD'                ~ 'GRAD',
      ENRL_ACAD_CAREER == 'UGRD'                ~ 'UGRD',
      TRUE                                      ~ 'Other'
    )
  )

stud_classified %>%
  count(CAREER_GROUP, ENRL_ACAD_CAREER, sort = TRUE)
```

## 4b — Graduate Students: Keep Graduate-Level Courses Only

Grad students sometimes take undergraduate courses. Only credit hours from
graduate-level classes count toward the GRAD billing bucket.

```{r grad-filter}
# Review academic group labels for GRAD students before filtering
stud_classified %>%
  filter(CAREER_GROUP == 'GRAD') %>%
  count(CLASS_ACAD_GROUP_LDESC, sort = TRUE)
```


```{r apply-grad-filter}
# Confirm 'Graduate' is the correct label in your data and update if needed
GRAD_COURSE_GROUPS <- c('Graduate')

stud_for_billing <- stud_classified %>%
  filter(
    !(CAREER_GROUP == 'GRAD' & !CLASS_ACAD_GROUP_LDESC %in% GRAD_COURSE_GROUPS),
    CAREER_GROUP != 'WI 65+ Audit'
  )

```

## 4c — Assign Residency Category

For **traditional students**, CDR `RES_STATUS` is the authoritative source.
`RES_STATUS = 'E'` (Midwest Exchange) is recoded to **Midwest Tuition Non-Resident**.

For **UFlex students**, `RES_STATUS` will be `NA` (not in CDR), so we fall
back to the RDS `RES_TUITION_RES_LDESC` field and apply the same Exchange recode
via text matching as a safety net.

```{r assign-residency}


stud_for_billing <- stud_for_billing %>%
  mutate(
    RESIDENCY_CATEGORY = case_when(

      # ── Traditional students: use CDR RES_STATUS ───────────────────────────
      !is_uflex & RES_STATUS == 'R' ~ 'Resident',
      !is_uflex & RES_STATUS == 'A' ~ 'Resident - Alumni Discount',
      !is_uflex & RES_STATUS == 'E' ~ 'Midwest Tuition Non-Resident',   # recoded from Exchange
      !is_uflex & RES_STATUS == 'L' ~ 'Midwest Tuition Non-Resident',
      !is_uflex & RES_STATUS == 'M' ~ 'Non-Resident - Michigan Reciprocity',
      !is_uflex & RES_STATUS == 'S' ~ 'Non-Resident - Minnesota Reciprocity',
      !is_uflex & RES_STATUS == 'W' ~ 'Non-Resident - Wisconsin Return Discount',
      !is_uflex & RES_STATUS == 'V' ~ 'Non-Resident - Veteran',
      !is_uflex & RES_STATUS == 'B' ~ 'Non-Resident - Alumni Discount',
      !is_uflex & RES_STATUS == 'N' ~ 'Non-Resident',
      !is_uflex & !is.na(RES_STATUS) ~ paste0('Non-Resident - Other (', RES_STATUS, ')'),

      # ── UFlex students: fall back to RDS residency description ─────────────
      is_uflex ~ case_when(
        str_detect(
          RES_TUITION_RES_LDESC,
          regex("midwest.*(exchange|compact)|exchange.*(midwest|compact)", ignore_case = TRUE)
        )                            ~ 'Midwest Tuition Non-Resident',
        !is.na(RES_TUITION_RES_LDESC) ~ RES_TUITION_RES_LDESC,
        TRUE                         ~ 'Unknown'
      ),

      TRUE ~ 'Unknown'
    )
  )

getOption("na.print")  # see what it's currently set to
options(na.print = "NA")  # reset to default

# Verify residency category distribution
stud_for_billing %>%
  count(CAREER_GROUP, RESIDENCY_CATEGORY, sort = TRUE) %>%
  as_tibble() %>%
  print(n = 40)

```

## 4d — Investigate Unknown Residency

Any row where `RESIDENCY_CATEGORY == 'Unknown'` signals a gap: either the
student has no CDR `RES_STATUS` (and is not UFlex), or the UFlex
`RES_TUITION_RES_LDESC` value didn't match any expected label.  The checks
below identify whether the problem is concentrated in a specific career group,
term, or raw residency value so it can be resolved upstream or recoded.

```{r residency-unknown}
unknown_res <- stud_for_billing %>%
  filter(RESIDENCY_CATEGORY == 'Unknown')

cat("Total 'Unknown' residency rows:", nrow(unknown_res), "\n\n")

# 1. Which career groups are affected?
cat("--- By CAREER_GROUP ---\n")
unknown_res %>%
  count(CAREER_GROUP, sort = TRUE) %>%
  print()

# 2. For traditional students (UGRD/GRAD): what raw CDR RES_STATUS value
#    is present (or missing) that wasn't mapped?
cat("\n--- Traditional students: raw RES_STATUS values ---\n")
unknown_res %>%
  filter(!is_uflex) %>%
  count(RES_STATUS, sort = TRUE) %>%
  print()

# 3. For UFlex students: what raw RDS residency description didn't match?
cat("\n--- UFlex students: raw RES_TUITION_RES_LDESC values ---\n")
unknown_res %>%
  filter(is_uflex) %>%
  count(RES_TUITION_RES_LDESC, sort = TRUE) %>%
  print()

# 4. Is Unknown concentrated in particular terms?
cat("\n--- By term ---\n")
unknown_res %>%
  count(CLASS_TERM_CD, CLASS_TERM_SDESC, CAREER_GROUP, sort = TRUE) %>%
  print()

# 5. Sample rows — inspect all residency-related raw fields side by side
cat("\n--- Sample rows (up to 20 distinct students) ---\n")
unknown_res %>%
  select(ENRL_EMPLID, CLASS_TERM_CD, CAREER_GROUP, CLASSIF,
         RES_STATUS, RES_TUITION_RES_LDESC, RESIDENCY_CATEGORY) %>%
  distinct() %>%
  slice_head(n = 20) %>%
  print()
```


*We should also examine when CLASSIF is NA as that is key information as well*

---

# Step 5 — Credit Hours Summaries by Career Group

Separate tables are produced for each billing bucket so they can be used
independently for tuition rate calculations.

Grain for all three: one row per **student × term × subject**.  A student
taking two sections of the same subject in the same term has their credit
hours summed into a single row.

```{r credit-hours-ugrd}
credit_hours_ugrd <- stud_for_billing %>%
  filter(CAREER_GROUP == 'UGRD') %>%
  group_by(
    ENRL_EMPLID,
    CLASS_TERM_ACAD_YEAR,
    CLASS_TERM_CD,
    CLASS_TERM_SDESC,
    ENRL_ACAD_CAREER,
    CAREER_GROUP,
    CLASSIF,
    RES_STATUS,
    RESIDENCY_CATEGORY,
    CTERM_FA_LOAD_CODE,
    CLASS_SUBJECT_LDESC,
    CLASS_ACAD_GROUP_LDESC,
    CLASS_ACAD_ORG_LDESC
  ) %>%
  summarise(
    n_classes          = n(),
    credit_hrs_taken   = sum(ENRL_UNITS_TAKEN,   na.rm = TRUE),
    credit_hrs_billing = sum(ENRL_UNITS_BILLING,  na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(CLASS_TERM_ACAD_YEAR, CLASS_TERM_SDESC, ENRL_EMPLID,
          CLASS_SUBJECT_LDESC)

cat("Undergraduate — rows:", nrow(credit_hours_ugrd),
    "| credit hrs taken:", sum(credit_hours_ugrd$credit_hrs_taken), "\n")
credit_hours_ugrd
```

```{r credit-hours-grad}
credit_hours_grad <- stud_for_billing %>%
  filter(CAREER_GROUP == 'GRAD') %>%
  group_by(
    ENRL_EMPLID,
    CLASS_TERM_ACAD_YEAR,
    CLASS_TERM_CD,
    CLASS_TERM_SDESC,
    ENRL_ACAD_CAREER,
    CAREER_GROUP,
    CLASSIF,
    RES_STATUS,
    RESIDENCY_CATEGORY,
    CTERM_FA_LOAD_CODE,
    CLASS_SUBJECT_LDESC,
    CLASS_ACAD_GROUP_LDESC,
    CLASS_ACAD_ORG_LDESC
  ) %>%
  summarise(
    n_classes          = n(),
    credit_hrs_taken   = sum(ENRL_UNITS_TAKEN,   na.rm = TRUE),
    credit_hrs_billing = sum(ENRL_UNITS_BILLING,  na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(CLASS_TERM_ACAD_YEAR, CLASS_TERM_SDESC, ENRL_EMPLID,
          CLASS_SUBJECT_LDESC)

cat("Graduate — rows:", nrow(credit_hours_grad),
    "| credit hrs taken:", sum(credit_hours_grad$credit_hrs_taken), "\n")
credit_hours_grad
```

```{r credit-hours-uflx}
credit_hours_uflx <- stud_for_billing %>%
  filter(CAREER_GROUP == 'UFLX') %>%
  group_by(
    ENRL_EMPLID,
    CLASS_TERM_ACAD_YEAR,
    CLASS_TERM_CD,
    CLASS_TERM_SDESC,
    ENRL_ACAD_CAREER,
    CAREER_GROUP,
    CLASSIF,
    RES_STATUS,
    RESIDENCY_CATEGORY,
    CTERM_FA_LOAD_CODE,
    CLASS_SUBJECT_LDESC,
    CLASS_ACAD_GROUP_LDESC,
    CLASS_ACAD_ORG_LDESC
  ) %>%
  summarise(
    n_classes          = n(),
    credit_hrs_taken   = sum(ENRL_UNITS_TAKEN,   na.rm = TRUE),
    credit_hrs_billing = sum(ENRL_UNITS_BILLING,  na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(CLASS_TERM_ACAD_YEAR, CLASS_TERM_SDESC, ENRL_EMPLID,
          CLASS_SUBJECT_LDESC)

cat("UFlex — rows:", nrow(credit_hours_uflx),
    "| credit hrs taken:", sum(credit_hours_uflx$credit_hrs_taken), "\n")
credit_hours_uflx
```

---

# Step 6 — WI 65+ Audit Summary (Separate, Not Billed)

```{r wi65-summary}
wi65_summary <- stud_classified %>%
  filter(CAREER_GROUP == 'WI 65+ Audit') %>%
  group_by(
    CLASS_TERM_ACAD_YEAR,
    CLASS_TERM_SDESC,
    CLASS_SUBJECT_LDESC
  ) %>%
  summarise(
    n_enrollments      = n(),
    credit_hrs_taken   = sum(ENRL_UNITS_TAKEN,  na.rm = TRUE),
    credit_hrs_billing = sum(ENRL_UNITS_BILLING, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(CLASS_TERM_ACAD_YEAR, CLASS_TERM_SDESC, CLASS_SUBJECT_LDESC)

wi65_summary
```

---

# Data Quality Checks

```{r quality-checks}
# Overall missingness
stud_enroll_full %>%
  summarise(
    n_rows              = n(),
    missing_career      = sum(is.na(ENRL_ACAD_CAREER)),
    missing_fa_load     = sum(is.na(CTERM_FA_LOAD_CODE)),
    missing_residency   = sum(is.na(RES_RESIDENCY)),
    missing_tuition_res = sum(is.na(RES_TUITION_RES_LDESC)),
    missing_subject     = sum(is.na(CLASS_SUBJECT_LDESC))
  )

# CDR join coverage (expect NA only for UFlex)
stud_classified %>%
  group_by(CAREER_GROUP) %>%
  summarise(
    n              = n(),
    missing_classif    = sum(is.na(CLASSIF)),
    missing_res_status = sum(is.na(RES_STATUS))
  )

# Career group counts
stud_classified %>%
  count(CAREER_GROUP, sort = TRUE)

# Any 'Other' career group rows to investigate
stud_classified %>%
  filter(CAREER_GROUP == 'Other') %>%
  count(ENRL_ACAD_CAREER, CLASS_TERM_CD, sort = TRUE)

# Residency distribution in billing data
stud_for_billing %>%
  count(CAREER_GROUP, RESIDENCY_CATEGORY, sort = TRUE)
```
