---
title: "Student Credit Hour By Program"
author: "Tony Daza"
date: "2026-01-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



**JUMP TO ALL IN ONE**

# Credit Hour by program for different student groups 
To calculate and predict income from tuition the office needs credit hours per program (subject). In addition, the students should be broken up by undergrad, graduate, and flex. With those further divided by resident status. 

Because we need Flex data this should be pulled from rds as well as the need to have the most up to date data (CDR is a year behind). 

CDR - V_Student_Analytical has the resident information we want (But would not have flex students)

RDS - REC_STUDENT_TERM_V & REC_STUDENT_TERM_FACT_V & REC_CLASS_DIM_V

REC_STUDENT_TERM_V = This has student information such as career(Grad vs UG), CTerm_TERM (4-digit code), Academic Year, load code

REC_STUDENT_TERM_FACT_V = The individual credit hours 

REC_CLASS_DIM_V = rows per class, CLASS_SUBJECT_LDESC (Subject is the grain we want for the final dataset) & course ID
- (Or do they want Class_ACAD_ORG as the descriptor for subject/program?)

REC_CRSE_CATALOG_DIM_V = CRSE_COURSE_TITLE_LONG (This might work for Flex)

Residency in RDS:
CC_RESIDENCY_DIM_V = Can match on EMPLID, RES_RESIDENCY & RES_RESIDENCY_LDESC


## Student Terms Key 


```{r}

# res_status <- dbGetQuery(
#   con_rds,
#   "
#   SELECT 
#       RES_EMPLID,
#       RES_ACAD_CAREER_SDESC,
#       RES_EFFECTIVE_TERM,
#       RES_RESIDENCY,
#       RES_RESIDENCY_LDESC,
#       RES_RESIDENCY_DT
#   FROM PKS_RDS_USR.CC_RESIDENCY_DIM_V
#       WHERE RES_INSTITUTION = 'UWPKS'
#       AND RES_EFFECTIVE_TERM IN('1257', '1261', '1263', '1265', '6506', '6507', '6508', '6509', '6510', '6511', '6512', '6601')
#   "
# )


# Terms: 
## Summer 2025-2026 = 1267
## Spring 2025-2026 = 1265
## Winter 2025-2026 = 1263
## Fall 2025-2026 = 1261
## Summer 2025-2026 = 1257



# Flex terms??
## March - May 2026 = 6603
## Feb - April 2026 = 6602
## Jan - March 2026 = 6601
## Dec - Feb 2026 = 6512
## Nov - Jan 2026 = 6511
## Oct - Dec 2025 = 6510
## Sept - Nov 2025 = 6509
## Aug - Oct 2025 = 6508
## Jul - Sept 2025 = 6507
## Jun - Aug 2025 = 6506
```


## List of courses for the selected terms

Lacks sufficient detail but it is a list of the courses but will be combined with enrollment data for credit hours using class_sid

```{r}

classes_25 <- dbGetQuery(
  con_rds,
  "
  SELECT
      A1.CLASS_SID,
      A1.CLASS_TERM_CD,
      A1.CLASS_TERM_SDESC,
      A1.CLASS_TERM_ACAD_YEAR,
      A1.CLASS_ACAD_GROUP_LDESC,
      A1.CLASS_ACAD_ORG_LDESC,
      A1.CLASS_SUBJECT_LDESC
  FROM PKS_RDS_USR.REC_CLASS_DIM_V A1
  WHERE A1.CLASS_INSTITUTION_CD = 'UWPKS'
      AND A1.CLASS_TERM_CD IN('1257', '1261', '1263', '1265', '6506', '6507', '6508', '6509', '6510', '6511', '6512', '6601')
  "
)




```

## List of courses with students for those term

REC_CLASS_DIM_V link with REC Enrollment using Class_sid

This results in a list of student enrollment data by class. We will use credits (units) as credit hours and use subject of class to get credit hours by subject. 

This currently lacks sufficient student information though. 


```{r}

stud_enroll_classes_25 <- dbGetQuery(
  con_rds,
  "
  SELECT
      e.CTERM_SID,
      e.CLASS_SID,
      e.ENRL_EMPLID,
      e.ENRL_ACAD_CAREER,
      e.ENRL_UNITS_TAKEN,
      e.ENRL_UNITS_BILLING,
      e.ENRL_EARNED_CREDIT
      c.CLASS_TERM_CD,
      c.CLASS_TERM_SDESC,
      c.CLASS_TERM_ACAD_YEAR,
      c.CLASS_ACAD_GROUP_LDESC,
      c.CLASS_ACAD_ORG_LDESC,
      c.CLASS_SUBJECT_LDESC
  FROM PKS_RDS_USR.REC_ENROLLMENT_FACT_V e 
  JOIN PKS_RDS_USR.REC_CLASS_DIM_V c
    ON c.CLASS_SID = e.CLASS_SID
  WHERE c.CLASS_INSTITUTION_CD = 'UWPKS'
      AND c.CLASS_TERM_CD IN('1257', '1261', '1263', '1265', '6506', '6507', '6508', '6509', '6510', '6511', '6512', '6601')
  "
)



#############

unique_values(stud_enroll_classes_25)

```


### List of CTERM IDs 
This prevents pulling the full student-term table and just gets the ones with class enrollment data 

```{r}


cterm_sids <- stud_enroll_classes_25 %>%
  distinct(CTERM_SID) %>%
  pull(CTERM_SID)

cterm_sid_sql <- paste(sprintf("'%s'", cterm_sids), collapse = ",")



```


## Pull student term data 

Just pulling the student term data from student enrollment IDs for 2025 terms

```{r}

stud_term_25 <- dbGetQuery(
  con_rds,
  sprintf(
    "
    SELECT
        CTERM_SID,
        CTERM_EMPLID,
        CTERM_CATEGORY_SDESC,
        CTERM_ACAD_YEAR,
        CTERM_TERM_CD,
        CTERM_TERM_SDESC,
        CTERM_ACAD_LOAD_SDESC,
        CTERM_ACAD_LEVEL_BOT_LDESC
    FROM PKS_RDS_USR.REC_STUDENT_TERM_DIM_V
    WHERE CTERM_INSTITUTION_CODE = 'UWPKS'
      AND CTERM_SID IN (%s)
    ",
    cterm_sid_sql
  )
)



```


## Pull Student Residency data 

This should pull just the student residency data from students that enrolled in classes in 2025, but we have to use their most recent residency status using effective term. This is because some students have residency from before the enrolled term that gets lost if we just pull on enrolled term. 

### List of enrolled students

```{r}

keys <- stud_enroll_classes_25 %>%
  distinct(ENRL_EMPLID, CLASS_TERM_CD) %>%
  mutate(
    ENRL_EMPLID = as.character(ENRL_EMPLID),
    CLASS_TERM_CD = as.character(CLASS_TERM_CD)
  )


```


### Residency plus key to extract student data 

This uses the most recent residency term data to asign the student a residency status. 

```{r}


key_rows_sql <- paste0(
  "SELECT '", keys$ENRL_EMPLID, "' AS EMPLID, '", keys$CLASS_TERM_CD, "' AS TERM_CD FROM dual",
  collapse = " UNION ALL "
)

res_asof_25 <- dbGetQuery(
  con_rds,
  sprintf(
    "
    WITH keylist AS (
      %s
    ),
    res_ranked AS (
      SELECT
        k.EMPLID,
        k.TERM_CD,
        r.RES_ACAD_CAREER_SDESC,
        r.RES_EFFECTIVE_TERM,
        r.RES_RESIDENCY,
        r.RES_RESIDENCY_LDESC,
        r.RES_RESIDENCY_DT,
        ROW_NUMBER() OVER (
          PARTITION BY k.EMPLID, k.TERM_CD
          ORDER BY r.RES_EFFECTIVE_TERM DESC, r.RES_RESIDENCY_DT DESC
        ) AS rn
      FROM keylist k
      LEFT JOIN PKS_RDS_USR.CC_RESIDENCY_DIM_V r
        ON r.RES_EMPLID = k.EMPLID
       AND r.RES_INSTITUTION = 'UWPKS'
       AND r.RES_EFFECTIVE_TERM <= k.TERM_CD
    )
    SELECT *
    FROM res_ranked
    WHERE rn = 1
    ",
    key_rows_sql
  )
)




```


### Combining student data with Residency data


REC_STUDENT_TERM_DIM_V (Student data on terms) with Residency data

CTERM_EMPLID = RES_EMPLID

We will pull residency status as well as load descriptor (CTERM_ACAD_LOAD_SDESC) and academic level (CTERM_ACAD_LEVEL_BOT_LDESC)

Might as well include term information as well. (CTERM_CATEGORY_SDESC) and (CTERM_ACAD_YEAR) and (CTERM_TERM_CD) as well as (CTERM_TERM_SDESC)

Of course I still need to filter on (CTERM_INSTITUTION_CODE = UWPKS)

This can be joined to student residency data through CTERM_EMPLID

It can be joined to the class info through ENRL_EMPLID = CTERM_EMPLID


```{r}

stud_enroll_full_25 <- stud_enroll_classes_25 %>%
  left_join(stud_term_25, by = "CTERM_SID") %>%
  left_join(
    res_asof_25,
    by = c("ENRL_EMPLID" = "EMPLID", "CLASS_TERM_CD" = "TERM_CD")
  )



```



# Combine lists to get a list of students with class info

Link stud_enroll_classes_25 with student data using ENRL_EMPLID = CTERM_EMPLID as well as CTERM_SID = CTERM_SID


Include term to differentiate classes by term

```{r}





stud_enroll_full_25 <- stud_enroll_classes_25 %>%
  left_join(
    stud_demo_25 %>%
      select(
        CTERM_SID,
        CTERM_EMPLID,
        CTERM_TERM_CD,
        CTERM_CATEGORY_SDESC,
        CTERM_ACAD_YEAR,
        CTERM_TERM_SDESC,
        CTERM_ACAD_LOAD_SDESC,
        CTERM_ACAD_LEVEL_BOT_LDESC,
        RES_ACAD_CAREER_SDESC,
        RES_RESIDENCY,
        RES_RESIDENCY_LDESC
      ),
    by = "CTERM_SID"  
  )




```


This should subsequently show a class row list of students enrolled per term with subject information.

That list will likely need to be pivoted by subject with indicator for student career and resident status.



# All in one


This is likely the most accurate. The reason being, prior to this the residency data was being pulled in first by effective date, but that was not the same as enrollment date, leading to missing data from residency. 

Downstream this led to missing data in the student demographics which was then passed down to the final dataset. 

To remedy this, the pull had to be done in RDS by matching residency status to the class enrollment data. Each student's enrollment was matched to their most recent residency term (Ideally equal, but if that was not present than the most recent effective residency date).

Subsequently the enrollment class data was joined to the student term data and the residency data. That resulted in a more complete dataset. 

This new rule means I am pulling the residency status that would have been in effect for that term of enrollment, or the most recent residency status on file for that student as of the term selected.

*Note* I should include student classification. (In CDR S = Over 60 i.e. they don't pay). Need to find the variable in RDS. Maybe REC_STDNT_GROUPS_DIM_V or REC_STDNT_GROUPS_DIM_ALL_V

```{r}

stud_enroll_full_25 <- dbGetQuery(
  con_rds,
  "
  WITH enroll AS (
    SELECT
      e.CTERM_SID,
      e.CLASS_SID,
      e.ENRL_EMPLID,
      e.ENRL_ACAD_CAREER,
      e.ENRL_UNITS_TAKEN,
      e.ENRL_UNITS_BILLING,
      e.ENRL_EARNED_CREDIT,
      e.ENRL_STATUS_SDESC,
      c.CLASS_TERM_CD,
      c.CLASS_TERM_SDESC,
      c.CLASS_TERM_ACAD_YEAR,
      c.CLASS_ACAD_GROUP_LDESC,
      c.CLASS_ACAD_ORG_LDESC,
      c.CLASS_SUBJECT_LDESC
    FROM PKS_RDS_USR.REC_ENROLLMENT_FACT_V e
    JOIN PKS_RDS_USR.REC_CLASS_DIM_V c
      ON c.CLASS_SID = e.CLASS_SID
    WHERE c.CLASS_INSTITUTION_CD = 'UWPKS'
      AND c.CLASS_TERM_CD IN('1257','1261','1263','1265',
                             '6506','6507','6508','6509',
                             '6510','6511','6512','6601')
  ),
  stud_term AS (
    SELECT
      st.CTERM_SID,
      st.CTERM_EMPLID,
      st.CTERM_CATEGORY_SDESC,
      st.CTERM_ACAD_YEAR,
      st.CTERM_TERM_CD,
      st.CTERM_TERM_SDESC,
      st.CTERM_ACAD_LOAD_SDESC,
      st.CTERM_ACAD_LEVEL_BOT_LDESC,
      st.CTERM_FA_LOAD_CODE
    FROM PKS_RDS_USR.REC_STUDENT_TERM_DIM_V st
    WHERE st.CTERM_INSTITUTION_CODE = 'UWPKS'
      AND st.CTERM_TERM_CD IN('1257','1261','1263','1265',
                              '6506','6507','6508','6509',
                              '6510','6511','6512','6601')
  ),
  res_ranked AS (
    SELECT
      en.ENRL_EMPLID,
      en.CLASS_TERM_CD,
      r.RES_ACAD_CAREER_SDESC,
      r.RES_EFFECTIVE_TERM,
      r.RES_RESIDENCY,
      r.RES_RESIDENCY_LDESC,
      r.RES_TUITION_RES_SDESC,
      r.RES_TUITION_RES_LDESC,
      r.RES_RESIDENCY_DT,
      ROW_NUMBER() OVER (
        PARTITION BY en.ENRL_EMPLID, en.CLASS_TERM_CD
        ORDER BY r.RES_EFFECTIVE_TERM DESC, r.RES_RESIDENCY_DT DESC
      ) AS rn
    FROM enroll en
    LEFT JOIN PKS_RDS_USR.CC_RESIDENCY_DIM_V r
      ON r.RES_EMPLID = en.ENRL_EMPLID
     AND r.RES_INSTITUTION = 'UWPKS'
     AND r.RES_EFFECTIVE_TERM <= en.CLASS_TERM_CD
  )
  SELECT
    en.*,
    st.CTERM_CATEGORY_SDESC,
    st.CTERM_ACAD_LOAD_SDESC,
    st.CTERM_ACAD_LEVEL_BOT_LDESC,
    st.CTERM_FA_LOAD_CODE,
    rr.RES_EFFECTIVE_TERM,
    rr.RES_RESIDENCY,
    rr.RES_RESIDENCY_LDESC,
    rr.RES_TUITION_RES_SDESC,
    rr.RES_TUITION_RES_LDESC,
    rr.RES_RESIDENCY_DT
  FROM enroll en
  LEFT JOIN stud_term st
    ON st.CTERM_SID = en.CTERM_SID
   AND st.CTERM_EMPLID = en.ENRL_EMPLID
   AND st.CTERM_TERM_CD = en.CLASS_TERM_CD
  LEFT JOIN res_ranked rr
    ON rr.ENRL_EMPLID = en.ENRL_EMPLID
   AND rr.CLASS_TERM_CD = en.CLASS_TERM_CD
   AND rr.rn = 1
  "
)




```


## Missingness 

None, must have pulled all the student residency data.

```{r}


# Spot check missingness
stud_enroll_full_25 %>%
  summarise(
    n_rows = n(),
    missing_student_term = sum(is.na(CTERM_CATEGORY_SDESC)),
    missing_residency = sum(is.na(RES_RESIDENCY))
  )

```

# Count 


Let's see if we group by term, the counts of student credit hours. 

```{r}


unique_values(stud_enroll_full_25, RES_TUITION_RES_LDESC)

unique_values(stud_enroll_full_25, CTERM_ACAD_LEVEL_BOT_LDESC)

unique_values(stud_enroll_full_25, CTERM_FA_LOAD_CODE)



stud_enroll_full_25 %>%
  group_by(CLASS_TERM_SDESC, CLASS_SUBJECT_LDESC, ENRL_ACAD_CAREER, RES_RESIDENCY) %>%
  count(ENRL_UNITS_TAKEN, ENRL_UNITS_BILLING)


# Enroll units taken and units billing are identical in this dataset. No differences



```


